Popularity Adjusted Block Models are Generalized Random Dot Product Graphs

### 2022-02-02 referee-response.Rmd

12. p10, line 12. $c^{(i)}$ is not defined.  
\textcolor{blue}{
$c^{(i)}$ is the $i^{th}$ entry of column vector $c$. 
This has been added to the paper.
}

13. p10. I didn't understand the exposition of the SSC algorithm.  
\textcolor{blue}{
We have added more detail to our description of the SSC algorithm. 
In particular, we describe the intuition/motivation behind the algorithm. 
}  
\textcolor{blue}{The description of SSC now reads:
}  
\textcolor{blue}{
``The SSC algorithm can be described as follows: 
Given \(X \in \mathbb{R}^{n \times d}\) with vectors
\(x_i^\top \in \mathbb{R}^d\) as rows of \(X\), the optimization problem
\(c_i = \arg\min_{c} \|c\|_1\) subject to \(x_i = X^\top c\) and
\(c^{(i)} = 0\), where $c^{(i)}$ is the $i^{th}$ entry of $c$, is solved for each \(i \in [n]\). 
The solutions are collected into matrix
\(C = \bigl[ c_1 \mid \cdots \mid c_n \bigr]^\top\) to
construct an affinity matrix \(B = |C| + |C^\top|\). If each \(x_i\) lies
exactly on one of \(K\) subspaces, \(B\) describes an undirected graph
consisting of {\em at least} \(K\) disjoint subgraphs, i.e., \(B_{ij} = 0\) if \(x_i, x_j\) lie on different subspaces. 
The intuition here is that vectors that lie on the same subspace can be described as linear combinations of each other, assuming the number of vectors in the subspace is greater than the dimensionality of the subspace. 
Then once sparsity is enforced, for each $c_i$, its $j^{th}$ element $c_i^{(j)}$ is zero if $x_j$ belongs to a subspace that doesn't contain $x_i$, resulting in $B_{ij} = 0$. 
Thus, for each $c_i$, $c_i^{(j)}$ is zero if $x_i$ and $x_j$ belong to different subspaces and may be nonzero if they do. 
If \(X\) instead represents points near \(K\) subspaces with some noise, 
then this property may only hold approximately and 
a final graph partitioning step may be required 
(e.g., edge thresholding or spectral clustering).''
}

14. p11, line 31. "... SSC on the spectral embedding of $A$". Could you support this assertion by references?  
\textcolor{blue}{
There are no references for this, as it is a novel idea. 
SSC has been proposed for community detection for the PABM but for the rows of $A$ rather than the ASE of $A$ \citep{noroozi2019estimation}. 
The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
}  
\textcolor{blue}{
We have added, ``... SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces.''
}

### 2022-02-02 referee-response.tex

  The motivation for our implementation is from Theorem 1, which says that the PABM is a GRDPG for which the latent positions are comprised of subspaces. 
  }\\
  \textcolor{blue}{
  We have added, ``... SSC on the rows of the spectral embedding of $A$, since the embedding of $P$ consists of $K$ subspaces (Theorem 1), and so the embedding of $A$ will lie approximately on the $K$ subspaces.''

### 2022-02-02 s675/blockmodels-and-rdpg.Rmd

@@

* To make this an assortative DCBM, set $p q > r^2$.
* In this example, $p = q = 1/4$, $r = 1/2$, and $\omega_i \in (0, 1)$.

@@

```{r, fig.height = 4, fig.width = 4, out.width = '100%'}
n1 <- 100
n2 <- 100
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/4
q <- 1/4
r <- 1/8
omega <- rbeta(n, 2, 1)
Omega <- tcrossprod(omega)
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
P <- P * Omega
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

@@

>>>>

# Two Types of Bernoulli Graphs

## Block Models

* $P_{ij} = f(z_i, z_j)$

* Community structure

* Estimation via EM
  * May require *ad hoc* approximations
  * Can be sensitive to initial guess

## (G)RDPGs

* $P_{ij} = x_i^\top I_{p,q} x_j$

* Latent space structure

* Estimation via ASE
  * Consistency theorem
  * Straightforward algorithm
  * Non-identifiable linear transformation

## All Bernoulli Graph models are (G)RDPGs

* Can we use this for inference on Block Models?

@@

* To make this an assortative DCBM, set $p q > r^2$.
* In this example, $p = q = 1/2$, $r = 1/8$, and $\omega_i \in (0, 1)$.

@@

```{r, fig.height = 4, fig.width = 4, out.width = '100%'}
n1 <- 100
n2 <- 100
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))
p <- 1/2
q <- 1/2
r <- 1/8
omega <- rbeta(n, 2, 1)
Omega <- tcrossprod(omega)
P <- matrix(r, nrow = n, ncol = n)
P[seq(n1), seq(n1)] <- p
P[seq(n1 + 1, n), seq(n1 + 1, n)] <- q
P <- P * Omega
A <- draw.graph(P)
qgraph::qgraph(A, vsize = 4, groups = factor(z), legend = FALSE)
```

@@

$$P_{ij} = \begin{cases} 
p \omega_i \omega_j & z_i, z_j = 1 \\
q \omega_i \omega_j & z_i, z_j = 2 \\
r \omega_i \omega_j & z_i \neq z_j
\end{cases}$$

### 2022-02-02 s675/rdpg.Rmd

\textcolor{darkred}{Random Dot Product Graphs}

@@

* For what types of graphs can we justify a latent structure?
* Any $A \sim \BG(P)$ is a RDPG if $P$ is positive semidefinite
* Whether we think of a graph as a RDPG depends on whether the latent structure is useful
* What about $P$ that are not positive semidefinite?

>>>>

\textcolor{darkred}{(Generalized) Random Dot Product Graphs}

@@

* For what types of graphs can we justify a latent structure?
* Any $A \sim \BG(P)$ is a RDPG if $P$ is positive semidefinite  
$P = V \Lambda V^\top \implies X = V \sqrt{\Lambda}$  
$X X^\top = V \sqrt{\Lambda} \sqrt{\Lambda} V^\top = V \Lambda V^\top = P$
* Whether we think of a graph as a RDPG depends on whether the latent structure is useful
* What about $P$ that are not positive semidefinite?
